---
title: "Kenya Data - First Look and Cleaning"
output: pdf_document
---

##Loading Data
Load in the data with new column headings:
```{r}
# Load the mapping table of new to old column names and make into a vector
ColumnNameData <- read.csv("NewColumnNameMapping.csv")
ColumnNameMap <- ColumnNameData$NewColName
names(ColumnNameMap) <- ColumnNameData$OldColName
#Make proposed names syntactically valid for R
ColumnNameMap <- make.names(ColumnNameMap, unique=TRUE)
#Load in data using new column names
Kenya <- read.csv("Addressing the -1 for DataKind - KENYA SURVEY - KMM UPDATES 22 JAN cleaned data - Sheet1.csv",
                  header = FALSE, stringsAsFactors = FALSE, skip = 2, 
                  col.names = ColumnNameMap)

#Also add a function to make proper case
properCase <- function(x) {
  lower <- tolower(x)
  s <- strsplit(lower, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}
```

Show the dimensions of the loaded data to make sure it was loaded ok (rows, columns)
```{r}
dim(Kenya)
```

##Analysing breakdown of gender and age
For Q10 on sex, 1: Male, 2: Female

For Q11 on age, 1: <=30, 2: 31-41, 3: 41-50, 4: 57-70, 5: >70
```{r}
table(Kenya$Q10.Sex)
table(Kenya$Q11.Age)
table(Kenya$Q10.Sex, Kenya$Q11.Age)
```

##Standardising text fields (Q3, 4 and 7)
Convert the 'Cooperative Society' field, Question 3, to proper case then make a table of all values.

(Question: Are any of these similar spellings actually the same?)
```{r}
Kenya$Q3.CoopSociety <- sapply(Kenya$Q3.CoopSociety, properCase)
table(Kenya$Q3.CoopSociety)
```

Convert the 'Village' field, Question 4, to proper case then make a table of all values.

(Question: Are any of these similarly-spelled villages actually the same?)
```{r}
Kenya$Q4.Village <- sapply(Kenya$Q4.Village, properCase)
table(Kenya$Q4.Village)
```

Convert the 'Producer organisation' field, Question 7, to proper case then make a table of all values.

(Question: Are any of these similarly-spelled orgs actallly the same?)
```{r}
Kenya$Q7.ProducerOrg <- sapply(Kenya$Q7.ProducerOrg, properCase)
table(Kenya$Q7.ProducerOrg)
```

##Columns with only one value
Identify which columns only have one unique value
```{r}
names(Filter(function(x)(length(unique(x))<2), Kenya))
```

##Duplicate columns
Check if Q155 values are identical to Q158 values
```{r}
table(Kenya$Q155.HowLongUsingIdea == Kenya$Q158.HowLongUsingIdea)
```
Eliminate one field from the dataset since they are identical
```{r}
Kenya$Q158.HowLongUsingIdea <- NULL
```

##Cleaning free-text number fields (Q38-40, 61-3, 69-72, 77-9, 113, 118)
Note: Look into Q73-76 since column descriptions appear duplicative to 69-71

Show unique values in Q38:
```{r}
unique(Kenya$Q38.CashCropEarnings)
```

Replace all non-digit characters with blank space, convert field to integer, then show unique values in Q38 and also plot histogram:
```{r}
Kenya$Q38.CashCropEarnings <- as.integer(gsub("[^0-9]", "", Kenya$Q38.CashCropEarnings))
unique(Kenya$Q38.CashCropEarnings)
hist(Kenya$Q38.CashCropEarnings)
```

Show unique values in Q39
```{r}
unique(Kenya$Q39.CashCropPx)
```

Replace all non-digit and non "." characters with blank space, convert field to numeric, then show unique values in Q39 and also plot histogram
```{r}
Kenya$Q39.CashCropPx <- as.numeric(str_extract(Kenya$Q39.CashCropPx, "\\d{1,}\\.?(\\d{1,})?"))
unique(Kenya$Q39.CashCropPx)
hist(Kenya$Q39.CashCropPx)
```

Show unique values in Q40
```{r}
unique(Kenya$Q40.CostOfProduction)
```

Replace all non-digit characters with blank space, convert field to integer, then show unique values in Q40 and also plot histogram:
```{r}
Kenya$Q40.CostOfProduction <- as.integer(gsub("[^0-9]", "", Kenya$Q40.CostOfProduction))
unique(Kenya$Q40.CostOfProduction)
hist(Kenya$Q40.CostOfProduction)
```

Show unique values in Q61, replace all non-digit characters with blank space, convert field to integer, then show resulting unique values 
```{r}
unique(Kenya$Q61.CropEarning)
Kenya$Q61.CropEarning <- as.integer(gsub("[^0-9]", "", Kenya$Q61.CropEarning))
unique(Kenya$Q61.CropEarning)
```

Show unique values in Q62
Note: Q62 appears to be all in different units (sack, kg, bundle, punch, crate). Best to avoid this for analysis at the moment.
```{r}
unique(Kenya$Q62.CropSalePx)
```

Show unique values in Q63, replace all non-digit characters with blank space, convert field to integer, then show resulting unique values 
```{r}
unique(Kenya$Q63.ProductionCost)
Kenya$Q63.ProductionCost <- as.integer(gsub("[^0-9]", "", Kenya$Q63.ProductionCost))
unique(Kenya$Q63.ProductionCost)
```

*NOTE:* Similar analyses and modifications are made for individual crop results in Questions 69-79

Show unique values in Q113
```{r}
unique(Kenya$Q113.CostOfChange)
```

Show unique values in Q118
```{r}
unique(Kenya$Q118.CostStarting)
```

##Longitude and Latitude values
First check how many values are present (not NA), then plot histogram showing distribution
```{r}
length(which(!is.na(Kenya$Latitude)))
length(which(!is.na(Kenya$Longitude)))
summary(Kenya$Latitude)
hist(Kenya$Latitude)
summary(Kenya$Longitude)
hist(Kenya$Longitude)
```
